---
title: "Homework 4"
author: "Jiayu Shi"
output: github_document
---

```{r}
## install.packages("caret")
library('manipulate')
library('MASS') ## for 'mcycle'
library('manipulate') ## for 'manipulate'
# Load the caret package

library(caret)
```



```{r}
mcycle
y <- mcycle$accel
x <- matrix(mcycle$times, length(mcycle$times), 1)

plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
```

# Randomly split the mcycle data into training (75%) and validation (25%) subsets.

```{r}
data(mcycle)
# Split the data into training and test sets
set.seed(123) # Set seed for reproducibility
trainIndex <- createDataPartition(mcycle$times, p = 0.75, list = FALSE)
train <- mcycle[trainIndex, ]
test <- mcycle[-trainIndex, ]
```

# Using the mcycle data, consider predicting the mean acceleration as a function of time. Use the Nadaraya-Watson method with the k-NN kernel function to create a series of prediction models by varying the tuning parameter over a sequence of values.


```{r}
## k-NN kernel function
## x  - n x p matrix of training inputs
## x0 - 1 x p input where to make prediction
## k  - number of nearest neighbors
kernel_k_nearest_neighbors <- function(x, x0, k=1) {
  ## compute distance betwen each x and x0
  z <- t(t(x) - x0)
  d <- sqrt(rowSums(z*z))

  ## initialize kernel weights to zero
  w <- rep(0, length(d))
  
  ## set weight to 1 for k nearest neighbors
  w[order(d)[1:k]] <- 1
  
  return(w)
}
```

```{r}
## Make predictions using the NW method
## y  - n x 1 vector of training outputs
## x  - n x p matrix of training inputs
## x0 - m x p matrix where to make predictions
## kern  - kernel function to use
## ... - arguments to pass to kernel function
nadaraya_watson <- function(y, x, x0, kern, ...) {
  k <- t(apply(x0, 1, function(x0_) {
    k_ <- kern(x, x0_, ...)
    k_/sum(k_)
  }))
  yhat <- drop(k %*% y)
  return(yhat)
}
```


## Use the Nadaraya-Watson method with the k-NN kernel function to create a series of prediction models by varying the tuning parameter over a sequence of values.

### k = 1
```{r}
x_plot <- matrix(seq(min(x),max(x),length.out=1000),1000,1)
plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
lines(x_plot,nadaraya_watson(train$accel, train$times, x_plot,kern = kernel_k_nearest_neighbors, k=1),col="#882255", lwd=2)
```



### k = 2
```{r}

plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
lines(x_plot,nadaraya_watson(train$accel, train$times, x_plot,kern = kernel_k_nearest_neighbors, k=2),col="#882255", lwd=2)
```

### k = 5
```{r}
plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
lines(x_plot,nadaraya_watson(train$accel, train$times, x_plot,kern = kernel_k_nearest_neighbors, k=5),col="#882255", lwd=2)
```

### k = 10
```{r}
plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
lines(x_plot,nadaraya_watson(train$accel, train$times, x_plot,kern = kernel_k_nearest_neighbors, k=10),col="#882255", lwd=2)
```

### k = 20
```{r}
plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
lines(x_plot,nadaraya_watson(train$accel, train$times, x_plot,kern = kernel_k_nearest_neighbors, k=20),col="#882255", lwd=2)
```


# With the squared-error loss function, compute and plot the training error, AIC, BIC, and validation error (using the validation data) as functions of the tuning parameter.

```{r}
## Compute effective df using NW method
## y  - n x 1 vector of training outputs
## x  - n x p matrix of training inputs
## kern  - kernel function to use
## ... - arguments to pass to kernel function
effective_df <- function(y, x, kern, ...) {
  y_hat <- nadaraya_watson(y, x, x,
    kern=kern, ...)
  sum(diag(attr(y_hat, 'k')))
}

## loss function
## y    - train/test y
## yhat - predictions at train/test x
loss_squared_error <- function(y, yhat)
  (y - yhat)^2

## test/train error
## y    - train/test y
## yhat - predictions at train/test x
## loss - loss function
error <- function(y, yhat, loss=loss_squared_error)
  mean(loss(y, yhat))

## AIC
## y    - training y
## yhat - predictions at training x
## d    - effective degrees of freedom
aic <- function(y, yhat, d)
  error(y, yhat) + 2/length(y)*d

## BIC
## y    - training y
## yhat - predictions at training x
## d    - effective degrees of freedom
bic <- function(y, yhat, d)
  error(y, yhat) + log(length(y))/length(y)*d
```

```{r}
edf = 0
for(i in 1:20){
  edf[i] = effective_df(train$accel, matrix(train$times), kern =  kernel_k_nearest_neighbors,k=i)
}

## training error
training_error_plot = 0
for (i in 1:20) {
  y_hat = nadaraya_watson(train$accel, train$times, matrix(train$times),kern = kernel_k_nearest_neighbors, k=i)
  training_error_plot[i] = error(train$accel,y_hat,loss=loss_squared_error)
}

## testing error
testing_error_plot = 0
for (i in 1:20) {
  y_hat = nadaraya_watson(train$accel, train$times, matrix(test$times),kern = kernel_k_nearest_neighbors, k=i)
  testing_error_plot[i] = error(test$accel,y_hat,loss=loss_squared_error)
}

## AIC
aic_plot = 0
for (i in 1:20) {
  y_hat = nadaraya_watson(train$accel, train$times, matrix(train$times),kern = kernel_k_nearest_neighbors, k=i)
  aic_plot[i] <- aic(train$accel,y_hat, edf[i])
}


## BIC
bic_plot = 0
for (i in 1:20) {
  y_hat = nadaraya_watson(train$accel, train$times, matrix(train$times),kern = kernel_k_nearest_neighbors, k=i)
  bic_plot[i] <- bic(train$accel,y_hat, edf[i])
}
```


```{r}
plot(1:20, testing_error_plot,ylim=c(0,1300), xlab="tuning parameter: K", ylab="AIC, BIC, and error")
points(1:20, bic_plot,col="green")
points(1:20, aic_plot,col="red")
points(1:20, training_error_plot,col="blue")

```


# For each value of the tuning parameter, Perform 5-fold cross-validation using the combined training and validation data. This results in 5 estimates of test error per tuning parameter value.

```{r}
## 5-fold cross-validation of knnreg model
## create five folds
set.seed(1985)
fold  <- createFolds(mcycle$accel, k=5)
print(fold)
sapply(fold, length)  ## not all the same length
```


```{r}
cvknnreg <- function(kNN = 10, flds=fold) {
  cverr <- rep(NA, length(flds))
  for(tst_idx in 1:length(flds)) { ## for each fold
    
    ## get training and testing data
    mcycle_trn <- mcycle[-flds[[tst_idx]],]
    mcycle_tst <- mcycle[ flds[[tst_idx]],]
    
    ## fit kNN model to training data
    knn_fit <- knnreg(accel ~ times,
                      k=kNN, data=mcycle)
    
    ## compute test error on testing data
    pre_tst <- predict(knn_fit, mcycle_tst)
    cverr[tst_idx] <- mean((mcycle_tst$Income - pre_tst)^2)
  }
  return(cverr)
}
## Compute 5-fold CV for kNN = 1:20
cverrs <- sapply(1:20, cvknnreg)
print(cverrs) ## rows are k-folds (1:5), cols are kNN (1:20)
cverrs_mean <- apply(cverrs, 2, mean)
cverrs_sd   <- apply(cverrs, 2, sd)
```

# Plot the CV-estimated test error (average of the five estimates from each fold) as a function of the tuning parameter. Add vertical line segments to the figure (using the segments function in R) that represent one “standard error” of the CV-estimated test error (standard deviation of the five estimates from each fold).


```{r}
## Plot the results of 5-fold CV for kNN = 1:20
plot(x=1:20, y=cverrs_mean, 
     ylim=c(200,1000),
     xlab="'k' in kNN", ylab="CV Estimate of Test Error")
segments(x0=1:20, x1=1:20,
         y0=cverrs_mean-cverrs_sd,
         y1=cverrs_mean+cverrs_sd)
best_idx <- which.min(cverrs_mean)
points(x=best_idx, y=cverrs_mean[best_idx], pch=20)
abline(h=cverrs_mean[best_idx] + cverrs_sd[best_idx], lty=3)

```



# Interpret the resulting figures and select a suitable value for the tuning parameter.

The least complex model within one standard error of the best is chosen. The suitable value for tuning parameter is 18.
















